Visualizing Project Evolution Through Abstract Syntax Tree Analysis

### 2016-07-06 paper/document.tex
Update paper with my TODOs.

% TODO: say all these other tools used file-based stuff:
% gource, github, code_swarm, others...?

@@

\begin{figure}[ht]
\centering
\includegraphics[width=\columnwidth]{antlr-density}
\caption{Density plot of type and file coverage for commits. The graph shows what percentage of type and file coverage the majority of commits have. One can see that the majority of commits have a type and file coverage around 0\% to 20\%. We can also see a difference between these measurements on the higher end of the graph. File coverage seems to have a group of commits that had a file coverage between 20\% and 50\% whereas type coverage have a similar looking grouping but between 40\% and 60\%.\label{fig:antlr4-density}}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\columnwidth]{tika-density}
\caption{Density of type and file coverage for commits. Here we can see that commits did not have much file coverage between 25\% and 75\%. On the other hand there were a lot of commits that have type coverage around 50\%.\label{fig:tika-density}}
\end{figure}

@@

To determine if type coverage yields different information than file coverage, we analyzed four open source Java projects: ANTLR4, Apache BookKeeper, Apache Curator and Apache Tika (Table~\ref{tab:project-summary}). Three of the four projects were collected from the Apache foundation, as they host a large number of well-curated open source Java projects. We counted the file coverage and type coverage on a commit-by-commit basis for each author. That is, each datum represents a commit with three dimensions: file coverage, type coverage, and its commit date. Data is sorted in ascending order of commit date. Table~\ref{tab:summary} lists a summary of the statistics measured.

@@

Having semantically aware evolutionary visualizations also gives users who are unfamiliar with the project at-a-glance information on the evolution of the project.

@@

@@

@@

>>>>

@@

\begin{figure}[ht]
\centering
\includegraphics[width=\columnwidth]{antlr-density}
\caption{R density plot of type and file coverage for commits. The graph shows what percentage of type and file coverage the majority of commits have. One can see that the majority of commits have a type and file coverage around 0\% to 20\%. We can also see a difference between these measurements on the higher end of the graph. File coverage seems to have a group of commits that had a file coverage between 20\% and 50\% whereas type coverage have a similar looking grouping but between 40\% and 60\%.\label{fig:antlr4-density}}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\columnwidth]{tika-density}
\caption{R density plot of type and file coverage for commits. Here we can see that commits did not have much file coverage between 25\% and 75\%. On the other hand there were a lot of commits that have type coverage around 50\%.\label{fig:tika-density}}
\end{figure}

% TODO: discuss density graphs more

@@

% Discuss "well-curated"

To determine if type coverage yields different information than file coverage, we analyzed four open source Java projects: ANTLR4, Apache BookKeeper, Apache Curator and Apache Tika (Table~\ref{tab:project-summary}). We chose ANTLR4 because it is a large Java project with a rich history of commits. ANTLR4 is also a member of the Qualitas Corpus~\cite{qualitas}, a curated collection of Java systems curated for empirical software engineering research. We then collected three smaller projects from the Apache foundation, as they host a large number of well-curated open source Java projects. These particular Apache projects were chosen for their clear development practices and available documentation. We counted the file coverage and type coverage on a commit-by-commit basis for each author. That is, each datum represents a commit with three dimensions: file coverage, type coverage, and its commit date. Data is sorted in ascending order of commit date. Table~\ref{tab:summary} lists a summary of the statistics measured.

@@

% Discuss user study

@@

% Discuss viz design in comparison to code_swarm

The timeline visualization presented in this paper (Section~\ref{sec:viz}) is intended to be a practical tool that enables a software developer's investigation of a project's evolution. Our timeline view can be seen as an evolution of simple visualizations such as GitHub's code frequency~\cite{github-graphs}. GitHub's graph is static, however it allows one to see an entire project's history at a glance. TypeV improves this by enabling one to see a particular \emph{type's} history at a glance. Given the greater density of information available, we made our timeline interactive such that a user is able to drill-down and explore the data at their own pace. Using interaction, TypeV not only shows what types have changed in project, but makes it possible to see \emph{why} a type has changed. We believe conventionality is a strong point in a developer's goal to make interesting discoveries in a project's evolution, but this has to be confirmed in user studies.

@@

A weakness of the timeline view described in this paper is that details about an author's activities are buried. As future work, the data extracted by TypeV can be used to create a visualization similar to Gource~\cite{gource} or code\_swarm~\cite{codeswarm}. These visualizations animate authors scurrying back-and-forth over files they edit; we proposed to animate authors' activities over types. The focal point would be a particular type or group of types. The type itself may be represented abstractly as a shape or its UML representation that grows and shrinks as its usage increases and wanes over time.

@@

% Acknowledgments

\section*{Acknowledgments}

The authors are grateful to the reviewers for their helpful suggestions on improving this work.

% TODO: acknowledge Craig?
